Topics in Algorithms

### 2018-01-10 project/paper.tex
Moved algs to bottom.

\section{Oracle with $O(\lg k)$ query time}
As aforementioned, binary search on the iteration sequence of the Thorup-Zwick oracle is performed to drive down the query time to $\lg k$. Starting with the iteration sequence $\mathcal{I}=1,\hdots,k-1$, the idea is to identify $r=O(\lg k)$ subsequences $(\mathcal{I}_1=\mathcal{I}) \supset \mathcal{I}_1 \supset \hdots \supset \mathcal{I}_r$, where for $2 \leq j \leq t$ we have $|\mathcal{I}_{j-1}| \geq 2|\mathcal{I}_j|$. We wish to maintain the invariant that $dist_k(u,v,i)$ applied to the first element of any subsequence $\mathcal{I}_j$ outputs the stretch $(2k-1)$ estimate in $|\mathcal{\mathcal{I}_j}|$ time.
\\
\begin{defi}
For vertices $u,v \in V$ an index $j \in \mathcal{I}$ is $(u,v)$-terminal if
\begin{enumerate}
\item $j = k-1$ or
\item $j < k-1$ is even and $p_j(u) \in B_v$ or $p_{j+1}(v) \in B_u$ 
\end{enumerate}
We note that $j$ attains a $(u,v)$ terminal index, the algorithm terminates in the current or next iteration. \label{def:uvterm}
\end{defi}
The definition of $(u,v)$-terminality captures the essence of when $dist_k(u,v)$ terminates, i.e. when $p_j(u) \in B_v$. This happens either when 1) $j=k-1$, in which case, trivially, $p_{k-1}(u) \in B_v$ or 2) when $j < k-1$ is even and $p_j(u) \in B_v$ or $p_{j+1}(v) \in B_u$ (in this case $dist_k(u,v)$ terminates either in iteration $j$ or $j+1$), see Alg. \ref{alg:distk}.
\\
\begin{defi}
A subsequence $\mathcal{I'}=i_1, \hdots, i_2$ of $\mathcal{I}$ is $(u,v)$-feasible if
\begin{enumerate}
\item $i_1$ is even and
\item $d_{G}(p_{i_1}(u),u) \leq i_1 \cdot d_G(u,v)$ and
\item $i_2$ is $(u,v)$-terminal
\end{enumerate}
\label{def:uvfeas}
\end{defi}
Furthermore, the definition of $(u,v)$-feasibility

@@

The recursion tree for $bdist_k$ for $u$ will have $\theta(k/\log k)$ children and $\theta(k/\log k)$ non-child nodes. Computing $j$ for each subsequence in $T$ will take $O(\log k)$ time per node, so total preprocessing time for $u$ is $O(k)$.  
\bibliographystyle{plain}
\bibliography{bibliography}

>>>>

\section{Oracle with $O(\lg k)$ query time}
As aforementioned, binary search on the iteration sequence of the Thorup-Zwick oracle is performed to drive down the query time to $\lg k$. Starting with the iteration sequence $\mathcal{I}=1,\hdots,k-1$, the idea is to identify $r=O(\lg k)$ subsequences $(\mathcal{I}_1=\mathcal{I}) \supset \mathcal{I}_1 \supset \hdots \supset \mathcal{I}_r$, where for $2 \leq j \leq t$ we have $|\mathcal{I}_{j-1}| \geq 2|\mathcal{I}_j|$. We wish to maintain the invariant that $dist_k(u,v,i)$ applied to the first element of any subsequence $\mathcal{I}_j$ outputs the stretch $(2k-1)$ estimate in $|\mathcal{I}_j|$ time.
\\
\begin{defi}
For vertices $u,v \in V$ an index $j \in \mathcal{I}$ is $(u,v)$-terminal if
\begin{enumerate}
\item $j = k-1$ or
\item $j < k-1$ is even and $p_j(u) \in B_v$ or $p_{j+1}(v) \in B_u$ 
\end{enumerate}
We note that $j$ attains a $(u,v)$ terminal index, the algorithm terminates in the current or next iteration. \label{def:uvterm}
\end{defi}
The definition of $(u,v)$-terminality captures the essence of when $dist_k(u,v)$ terminates, i.e. when $p_j(u) \in B_v$. This happens either when 1) $j=k-1$, in which case, trivially, $p_{k-1}(u) \in B_v$ or 2) when $j < k-1$ is even and $p_j(u) \in B_v$ or $p_{j+1}(v) \in B_u$ (in this case $dist_k(u,v)$ terminates either in iteration $j$ or $j+1$), see Alg. \ref{alg:distk}.
\\
\begin{defi}
A subsequence $\mathcal{I'}=i_1, \hdots, i_2$ of $\mathcal{I}$ is $(u,v)$-feasible if
\begin{enumerate}
\item $i_1$ is even and
\item $d_{G}(p_{i_1}(u),u) \leq i_1 \cdot d_G(u,v)$ and
\item $i_2$ is $(u,v)$-terminal
\end{enumerate}
\label{def:uvfeas}
\end{defi}
Condition $1$ and $3$ of $(u,v)$-feasibility entails a symmetric view of $dist_k(u,v,i)$ in which $u$ and $v$ successively exhanges roles for even an uneven $j$. Condition $2$ ensures that $dist_k(u,v,i)$ returns a $(2k-1)$-approximation when run with the first  $(u,v)$-feasible subsequence, as seen in the following lemma:

@@

The recursion tree for $bdist_k$ for $u$ will have $\theta(k/\log k)$ children and $\theta(k/\log k)$ non-child nodes. Computing $j$ for each subsequence in $T$ will take $O(\log k)$ time per node, so total preprocessing time for $u$ is $O(k)$.

\bibliographystyle{plain}
\bibliography{bibliography}
